*Prochlorococcus marinus* responses to light and oxygen

### 2024-02-28 Code/Manuscript_O2.Rmd
edits to text

---
title: "*Prochlorococcus marinus* responses to light and oxygen" 
author:
  - name: 'Mireille Savoie'
    institute: mta
    main: true
    email: msavoie@mta.ca
  - name: Aurora Mattison
    institute: [mta, ubc]
  - name: Julie Nadeau
    institute: mta
  - name: Laurel Genge
    institute: mta
  - name: Sylwia Śliwińska-Wilczewska
    institute: [mta, ug]
  - name: Maximilian Berthold
    institute: mta
  - name: Naaman M. Omar
    institute: mta
  - name: Ondřej Prášil
    institute: [mta, algatech]
    main: false
  - name: Amanda M. Cockshutt
    institute: [mta, sfx]
    main: false
  - name: Douglas A. Campbell
    institute: mta
    affil: 1
    main: false
    orcid: '0000-0001-8996-5463'
    email: 'dcampbel@mta.ca'
affiliation:
  - num: 1
    address: Department of Biology, Mount Allison University
institute:
  - mta: 'Department of Biology, Mount Allison University, 63B York St., Sackville NB, Canada, E4L 1G7.'
  - ubc: 'Department of Biology, University of British Columbia'
  - sfx: 'Department of Chemistry, St. Frances Xavier University'
  - algatech: 'Institute of Microbiology, Center Algatech, Laboratory of Photosynthesis, Novohradska 237, Trebon, CZ 37981, Czech Republic.'
  - ug: 'Institute of Oceanography, University of Gdansk, 46 Piłsudskiego St, Gdynia, Poland'
date: "`r format(Sys.Date())`"
output:
  bookdown::html_document2:
    code_folding: show
    keep_md: yes
    toc: TRUE
    toc_float: TRUE
    toc_depth: 6
    fig_caption: yes
    pandoc_args: 
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
  bookdown::word_document2:
    code_folding: show
    reference_docx: KnitTemplate.docx
    keep_md: yes
    toc: FALSE
    toc_depth: 6
    fig_caption: yes
    pandoc_args: 
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
always_allow_html: true
bibliography: [Manuscript_O2.bib, RPackages.bib]
csl: plos-one.csl
editor_options: 
  markdown: 
    wrap: 72
---

@@

## *Prochlorococcus marinus* diversity {.unnumbered}
*Prochlorococcus*, a genus of Cyanobacteria, is the smallest known photosynthetic prokaryote, with cell diameters ranging from 0.5 to 0.7 µm [@chisholmProchlorococcusMarinusNov1992]. Despite small cell size, *P. marinus* contribute 13 to 48% of net primary production in oligotrophic oceans, corresponding to about 30% of global oxygen production [@partenskyProchlorococcusMarinePhotosynthetic1999]. *Prochlorococcus marinus* growth is currently limited to between latitudes of 40°N to 40°S in open ocean waters, from surface to 300 m depth, thus spanning 3 orders of magnitudes of light [@partenskyProchlorococcusMarinePhotosynthetic1999; @chisholmProchlorococcusMarinusNov1992 ].

*Prochlorococcus marinus* is functionally differentiated from other cyanobacteria primarily by the absence of phycobilisomes [@chisholmProchlorococcusMarinusNov1992]. Instead, *Prochlorococcus*  uses intra-membrane Prochlorophyte Chlorophyll Binding (Pcb) proteins binding divinylchlorophyll *a* and *b*, and α-carotene, as major light harvesting complexes [@goerickePigmentsProchlorococcusMarinus1992]; [@larocheIndependentEvolutionProchlorophyte1996], with diversity in the number of genes encoding Pcb, expression of Pcb under changing light [@garczarekMultiplicationAntennaGenes2000; @rocapGenomeDivergenceTwo2003], and in thylakoid organization [@partenskyProchlorococcusMarinePhotosynthetic1999].  The genes encoding much of the photosynthetic apparatus are highly conserved in *Prochlorococcus* [@kettlerPatternsImplicationsGene2007], though the photosynthetic efficiency may vary among clades [@hessPhotosyntheticApparatusProchlorococcus2001; @moorePhotophysiologyMarineCyanobacterium1999].

*Prochlorococcus marinus* comprises many strains, organized into clades, defined by 16S-23S intergenic transcribed ribosomal sequence signatures [@rocapGenomeDivergenceTwo2003]. The clades inhabit distinct ecological niches [@moorePhysiologyMolecularPhylogeny1998], originally defined by Low-Light (LL) or High-Light (HL).  Of at least 12 known *Prochlorococcus* genetic clades only 5 to date have cultured representatives; HLI, HLII, LLI, LLII/III and LIV [@billerProchlorococcusStructureFunction2015].

Current niches of *P. marinus* strains span ocean water columns [ @veldhuisVerticalDistributionPigment1990; @chisholmNovelFreelivingProchlorophyte1988; @partenskyProchlorococcusMarinePhotosynthetic1999] and extend into regions with low dissolved oxygen concentrations [@partenskyProchlorococcusAdvantagesLimits2010; @goerickeNovelNicheProchlorococcus2000; @lavinNovelLineagesProchlorococcus2010, @ulloaCyanobacteriumProchlorococcusHas2021; @garcia-robledoCrypticOxygenCycling2017]. 
Low-Light clades thrive in deeper ocean waters, extending beyond 200 m in depth [@partenskyProchlorococcusMarinePhotosynthetic1999], where only ~1% of the surface irradiance penetrates, primarily in the blue (450 nm) to green (520 nm) spectral range [@holtropVibrationalModesWater2021]. Clade LLIV, including cultured strain MIT9313, falls near the base of the *Prochlorococcus* radiation, and is characterized by preference for low light, typical of depths from 120 m to 200 m [@partenskyProchlorococcusMarinePhotosynthetic1999]. LLIV members are, as yet, the only cultured strains to have been found in Oxygen Minimum Zones (OMZ).  Some as yet uncultured *P. marinus* strains in clades LLV and LLVI also thrive in OMZ of the subtropical Atlantic and Pacific Oceans, where dissolved oxygen concentrations [O~2~] can be less than 20 µM [@ulloaCyanobacteriumProchlorococcusHas2021; @lavinNovelLineagesProchlorococcus2010; @goerickeNovelNicheProchlorococcus2000; @garcia-robledoCrypticOxygenCycling2017; @johnsonEnergeticsGrowthKinetics1999]. *Prochlorococcus marinus* LL ecotypes may dominate the phytoplankton within OMZ [@lavinNovelLineagesProchlorococcus2010; @ulloaCyanobacteriumProchlorococcusHas2021; @partenskyProchlorococcusAdvantagesLimits2010], where they may be net O~2~ consumers [@partenskyComparisonPhotosyntheticPerformances2018]. Clade LLI includes cultured strain NATL2A, which prefers moderate irradiance typically between 30 and 100 m depth. Clades LLII and LLIII, including cultured strain SS120, are grouped together as the second oldest phylogenetic lineage diversifying in the *P. marinus* radiation, with an affinity for low light.

@@

Climate change is also rapidly changing ocean chemistry.  By the end of this century, surface ocean pH is projected to decline by 0.1 to 0.4  due to projected increases in carbon dioxide concentrations [@garcia-sotoOverviewOceanClimate2021]. Moreover, substantial changes in the global water cycle, leading to extensive changes in worldwide precipitation patterns, are affecting ocean salinity levels on a global scale, and ice melts due to rising temperatures are impacting salinity levels in the Arctic and Northwest Atlantic oceans [@leeRecognizingSalinityThreats2022]. Increasing sea temperatures are also causing decreases in [O~2~] across global oceans [@matearLongtermChangesDissolved2003], particularly toward the poles [@helmObservedDecreasesOxygen2011]. Warmer ocean waters increase stratification, and decrease oxygen solubility at the surface, which in turn decreases oxygen mixing downwards by ocean currents [@garcia-sotoOverviewOceanClimate2021]. Models predict that OMZ in the Pacific and Indian Oceans are expanding [@garcia-sotoOverviewOceanClimate2021; @buseckeDivergingFatesPacific2022]. In spite of these expanding OMZ @buseckeDivergingFatesPacific2022 highlight that the cores of the OMZ, where the oxygen levels are lowest, are actually projected to contract.

We used the OceansMap Protein Portal (OPP; https://proteinportal.whoi.edu/) [@saitoDevelopmentOceanProtein2021] to analyze the distribution of proteins from clades of *P. marinus* in samples tanken across a range of [O~2~] and depth, which in turn correlates to peak light at the site of sampling. We therefore analyzed the growth and physiological responses of representative strains from three clades of *P. marinus* under a matrix of [O~2~], light levels, spectral waveband ranges, and photoperiods, to approximate eco-physiological conditions representative of current and hypothetical future ocean zones. *Prochlorococcus marinus* MED4, a clade HLI strain, was isolated near the ocean surface (5m depth) of the Mediterranean Sea where [O~2~] is near saturation, light levels are high and spectral bias from full solar irradiance is minimal. *Prochlorococcus marinus* SS120, a clade LLIII strain, was isolated from the Sargasso Sea at a depth of 120 m, while *P. marinus* MIT9313, a clade LLIV strain, was isolated from the North Atlantic Gulf Stream at a depth of 135 m. At these depths, light attenuation and spectral shifts occur, resulting in low blue and green light, while [O~2~] varies from near-surface saturation levels to decreased concentrations, but does not necessarily decrease systematically with depth [@billheimerOxygenSeasonalityUtilization2021].

We then analyzed growth rates in terms of cumulative diel Photosynthetically Usable Radiation (PUR). Photosynthetic organisms absorb light energy within the Photosynthetically Active Radiation (PAR) range, 350 to 700 nm, for photosynthesis [@morelAvailableUsableStored1978]. PUR represents the fraction of PAR that is absorbed by the pigments of a given photosynthetic organism [@morelAvailableUsableStored1978], taking into account the specific spectral wavebands these pigments absorb. *Prochlorococcus marinus* light-harvesting complexes, [@chisholmProchlorococcusMarinusNov1992;
@goerickePigmentsProchlorococcusMarinus1992], show an absorption maxima
of 442 nm for divinyl chlorophyll a and 478 nm for divinyl chlorophyll b
[@goerickePigmentsProchlorococcusMarinus1992] allowing *P. marinus* to
efficiently harvest blue light in the 400 nm
to 500 nm range [@morelAvailableUsableStored1978] prevailing in deep ocean
habitats, where only blue spectral wavelengths prevail [@holtropVibrationalModesWater2021].  In *P. marinus* small cell diameters, from 0.5 to 0.7 µm [@chisholmProchlorococcusMarinusNov1992], and simple cell structures, minimize the complication of pigment package effect or intracellular self-shading [@morelProchlorococcusSynechococcusComparative1993]. Given the different spectral light regimes typical of the niches of different ecotypes, expressing growth rates in terms of cumulative diel PUR might simplify different photoperiods, spectral bands, and PAR levels into a common parameter, making growth response comparisons across strains and different oxygen levels more accessible.  We also aimed to detect whether growth responses are driven simply by cumulative diel PUR, or whether specific photoperiods, spectral bands or PAR levels have independent, albeit interacting, effects on growth.

@@

Collection and treatment of protein samples were performed following Saito et al., [@saitoMultipleNutrientStresses2014; @saitoNeedlesBlueSea2015]. Briefly, samples of seawater from depths of 20, 40, 50, 60, 70, 80, 90, 120, 150, 200, 250, 300, 380, 400, 500, 550, 600, and 800 m below the ocean surface were pumped through a 0.2 µm filter, preserved in RNAlater and frozen at -80°C until extraction. Proteins were extracted from the filter in an SDS-based detergent, embedded in tube gel, alkylated and reduced prior to in-gel trypsin digestion. Peptide spectra were generated using a Q-Exactive Orbitrap Mass Spectrometer, searched in the SEQUEST CITATIONXXXX and labelled with the most likely protein and species annotation from Uniprot.  Oxygen levels at the location of sampling were recorded.

@@

```{r ProchloroPhotosynthDepthO2, fig.cap = ProchloroPhotosynthDepthO2_cap , fig.height = 20, fig.width = 10, echo = FALSE}

ProchloroPhotosynthDepthO2_cap <- glue("**Ocean detection of *Prochlorococcus marinus* photosynthesis complexes.**", 
                       "Protein detections are plotted vs. O~2~ (µM) (X axis) and depth (m) (Y axis) at sample origin.",
                       "Rows separate data annotated as from *Prochlorococcus marinus* strains MED4 (Clade HLI), SS120 (Clade LLIII) and MIT9313 (Clade LLIV).",
                       "Columns show detections of proteins annotated as  Photosystem II (PSII), Cytochromeb6f complex (Cytb6f), Photosystem I (PSI), ATP Synthase or Ribulose-1,5-bisphosphate oxygenase carboxylase (RUBISCO).",
                       "For comparison culture growth experimental conditions are indicated by horizontal grey lines for depths approximating imposed peak Photosynthetically Active Radiation (µmol photons m^-2^ s^-1^); and vertical grey lines for [O~2~] (µM).",
                       "Point colours identify *Prochlorococcus* clade assignments",
                       "Data from OceanProteinPortal (https://www.oceanproteinportal.org/).",
 .sep = " ")

knitr::include_graphics(file.path(FiguresIn, "ProchloroPhotosynthDepthO2.png"))
```

@@

 BluevsRedPurFitsPlots_cap <- glue("**Chlorophyll specific growth rate (d^-1^) vs. cumulative diel Photosynthetic Usable Radiation (PUR, µmol photons m^-2^ d^-1^).**",
                    "Rows show levels of imposed dissolved O~2~ concentrations as 250 µM, 25 µM and 2.5 µM.",
                    "Columns are strains; MED4, SS120 and MIT9313.",
                    "Shapes show the imposed photoperiod (h); 4 h (solid square),  8 h (solid diamond), 12 h (solid circle), 16 h (solid upright triangle).",
                      "Symbol colours show the spectral waveband for growth; 660 nm (red symbols), and 450 nm (blue symbols).",
                      "Large symbols show mean of growth rate from logistic curve fits; small symbols show values for replicate determinations, if any.",
                    "Platt (add Platt citation here) 4 parameter model fit to 660 nm (red lines) and 450 nm (blue lines) growth data for each combination of strain and dissolved oxygen shown with solid lines (significantly different fits, pvalue < 0.05) or dashed lines (not significantly different fits, pvalue > 0.05).",
                       .sep = " ")

@@

```{r ProchloroProteinMetabDepthO2, fig.cap = ProchloroProteinMetabDepthO2_cap , fig.height = 20, fig.width = 10, echo = FALSE}

ProchloroProteinMetabDepthO2_cap <- glue("**Ocean detection of *Prochlorococcus marinus* protein metabolism complexes.**", 
                       "Protein detections are plotted vs. O~2~ (µM) (X axis) and depth (m) (Y axis) at sample origin.",
                       "Rows separate data annotated as from *Prochlorococcus marinus* strains MED4 (Clade HLI), SS120 (Clade LLIII) and MIT9313 (Clade LLIV).",
                       "Columns show detections of proteins annotated as FtsH Protease Complexes (FtsH1, FtsH2, FtsH3) or the Ribosome).",
                       "Culture growth experimental conditions indicated by horizontal grey lines for depths approximating Photosynthetically Active Radiation (µmol photons m^-2^ s^-1^) and vertical grey lines for [O~2~] (µM).",
                       "Point colours identify *Prochlorococcus* clade assignments",
                       "Data from OceanProteinPortal (https://www.oceanproteinportal.org/).",
 .sep = " ")

knitr::include_graphics(file.path(FiguresIn, "ProchloroProteinMetabDepthO2.png"))
```

@@

*Prochlorococcus marinus* MED4 has a physiological requirement for more than 4 h of light per day, and thus this strain will not exploit habitats with short photoperiods, typical of winter or light attenuated depths. MED4 is also excluded from the lowest oxygen habitats, equivalent to our 2.5 µM O~2~ experimental conditions. MED4 can, however, grow under OMZ regions with slightly higher [O~2~], as by the growth under our 25 µM O~2~ experiments. Genomic analyses  ( Figure \@ref(fig:ProchlorococcusEnzymeKms))  and previous transcriptional analyses [@bergResponsesPsbAHli2011]  suggests MED4 is excluded from growth below ~ 25 µM O~2~ because it relies upon a ubiquinol oxidase, non-electrogenic, to maintain oxidation/reduction balance in the intersystem electron transport chain.  On the other hand, MED4 shows inducible expression of FtsH isoforms [@bonisteelStrainSpecificDifferences2018], to counter photoinactivation of PSII under higher PAR and [O~2] environments. Photoinactivation, does, however, impose an increased cost of growth upon MED4, since growth under red light, to lower the rate of photoinactivation of PSII [@murphyPhotoinactivationPhotosystemII2017], allows MED4 to achieve faster growth per absorbed photon than growth under red light. TARA Oceans
Project data [@delmontLinkingPangenomesMetagenomes2018] indeed reported presence of *P. marinus* MED4-like genomes at depths ranging from 5 m to 90 m, representing high to low blue light levels, in the Pacific South East Ocean. @delmontLinkingPangenomesMetagenomes2018 did not analyze data from depths beyond the subsurface chlorophyll maximum layer, the layer in the water column where the chlorophyll a concentration peaks, nor did they report the [O~2~] at those depths. Our growth findings are also consistent with Figure\@ref(fig:ProchloroPhotosynthDepthO2) showing PSII proteins annotated as MED4, at depths up to 200 meters with O~2~ of 15 µM.

>>>>

---
title: "*Prochlorococcus marinus* responses to light and oxygen" 
author:
  - name: 'Mireille Savoie'
    institute: mta
    main: true
    email: msavoie@mta.ca
  - name: Aurora Mattison
    institute: [mta, ubc]
  - name: Julie Nadeau
    institute: mta
  - name: Laurel Genge
    institute: mta
  - name: Sylwia Śliwińska-Wilczewska
    institute: [mta, ug]
  - name: Maximilian Berthold
    institute: mta
  - name: Naaman M. Omar
    institute: mta
  - name: Ondřej Prášil
    institute: [mta, algatech]
    main: false
  - name: Amanda M. Cockshutt
    institute: [mta, sfx]
    main: false
  - name: Douglas A. Campbell
    institute: mta
    affil: 1
    main: false
    orcid: '0000-0001-8996-5463'
    email: 'dcampbel@mta.ca'
affiliation:
  - num: 1
    address: Departments of Biology and Chemistry and Biochemistry, Mount Allison University
institute:
  - mta: 'Department of Biology, Mount Allison University, 63B York St., Sackville NB, Canada, E4L 1G7.'
  - ubc: 'Department of Biology, University of British Columbia'
  - sfx: 'Department of Chemistry, St. Frances Xavier University'
  - algatech: 'Institute of Microbiology, Center Algatech, Laboratory of Photosynthesis, Novohradska 237, Trebon, CZ 37981, Czech Republic.'
  - ug: 'Institute of Oceanography, University of Gdansk, 46 Piłsudskiego St, Gdynia, Poland'
date: "`r format(Sys.Date())`"
output:
  bookdown::html_document2:
    code_folding: show
    keep_md: yes
    toc: TRUE
    toc_float: TRUE
    toc_depth: 6
    fig_caption: yes
    pandoc_args: 
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
  bookdown::word_document2:
    code_folding: show
    reference_docx: KnitTemplate.docx
    keep_md: yes
    toc: FALSE
    toc_depth: 6
    fig_caption: yes
    pandoc_args: 
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
always_allow_html: true
bibliography: [Manuscript_O2.bib, RPackages.bib]
csl: plos-one.csl
editor_options: 
  markdown: 
    wrap: 72
---

@@

## *Prochlorococcus marinus* diversity {.unnumbered}
*Prochlorococcus*, a genus of Cyanobacteria, is the smallest known photosynthetic prokaryote, with cell diameters ranging from 0.5 to 0.7 µm [@chisholmProchlorococcusMarinusNov1992]. Despite small cell size, *P. marinus* contribute 13 to 48% of net primary production in oligotrophic oceans, corresponding to about 30% of global oxygen production [@partenskyProchlorococcusMarinePhotosynthetic1999]. *Prochlorococcus marinus* growth is currently limited to between latitudes of 40°N to 40°S in open ocean waters, from surface to 300 m depth, thus spanning 3 orders of magnitudes of light irradiance [@partenskyProchlorococcusMarinePhotosynthetic1999; @chisholmProchlorococcusMarinusNov1992 ].

*Prochlorococcus marinus* is functionally differentiated from other cyanobacteria primarily by the absence of phycobilisomes [@chisholmProchlorococcusMarinusNov1992]. Instead, *Prochlorococcus*  uses intra-membrane Prochlorophyte chlorophyll binding (Pcb) proteins binding divinylchlorophyll *a* and *b*, and α-carotene, as major light harvesting complexes [@goerickePigmentsProchlorococcusMarinus1992]; [@larocheIndependentEvolutionProchlorophyte1996], with diversity in the number of genes encoding Pcb, expression of Pcb under changing light [@garczarekMultiplicationAntennaGenes2000; @rocapGenomeDivergenceTwo2003], and in thylakoid organization [@partenskyProchlorococcusMarinePhotosynthetic1999].  The genes encoding much of the photosynthetic apparatus are highly conserved in *Prochlorococcus* [@kettlerPatternsImplicationsGene2007], though the photosynthetic efficiency may vary among clades [@hessPhotosyntheticApparatusProchlorococcus2001; @moorePhotophysiologyMarineCyanobacterium1999].

*Prochlorococcus marinus* comprises many strains, organized into clades, defined by 16S-23S intergenic transcribed ribosomal sequence signatures [@rocapGenomeDivergenceTwo2003]. The clades inhabit distinct ecological niches [@moorePhysiologyMolecularPhylogeny1998], originally defined by High-Light (HL) or Low-Light (LL).  Of at least 12 known *Prochlorococcus* genetic clades only 5 to date have cultured representatives; HLI, HLII, LLI, LLII/III and LIV [@billerProchlorococcusStructureFunction2015].

Current niches of *P. marinus* strains span ocean water columns [@veldhuisVerticalDistributionPigment1990; @chisholmNovelFreelivingProchlorophyte1988; @partenskyProchlorococcusMarinePhotosynthetic1999] and extend into regions with low dissolved oxygen concentrations [@partenskyProchlorococcusAdvantagesLimits2010; @goerickeNovelNicheProchlorococcus2000; @lavinNovelLineagesProchlorococcus2010, @ulloaCyanobacteriumProchlorococcusHas2021; @garcia-robledoCrypticOxygenCycling2017].

Low-Light clades thrive in deeper ocean waters, extending beyond 200 m in depth [@partenskyProchlorococcusMarinePhotosynthetic1999], where only ~1% of the surface irradiance penetrates, primarily in the blue (450 nm) to green (520 nm) spectral range [@holtropVibrationalModesWater2021]. Clade LLI includes cultured strain NATL2A, which prefers moderate irradiance typically between 30 and 100 m depth. Clades LLII and LLIII, including cultured strain SS120, are grouped together as the second oldest phylogenetic lineage diversifying in the *P. marinus* radiation, with an affinity for low light. Clade LLIV, including cultured strain MIT9313, falls near the base of the *Prochlorococcus* radiation, and is characterized by preference for low light, typical of depths from 120 m to 200 m [@partenskyProchlorococcusMarinePhotosynthetic1999]. LLIV members are, as yet, the only cultured strains to have been found in Oxygen Minimum Zones (OMZ).  Some, as yet, uncultured *P. marinus* strains in clades LLV and LLVI also thrive in OMZ of the subtropical Atlantic and Pacific Oceans, where dissolved oxygen concentrations [O~2~] can be less than 20 µM [@ulloaCyanobacteriumProchlorococcusHas2021; @lavinNovelLineagesProchlorococcus2010; @goerickeNovelNicheProchlorococcus2000; @garcia-robledoCrypticOxygenCycling2017; @johnsonEnergeticsGrowthKinetics1999]. *Prochlorococcus marinus* LL ecotypes may dominate the phytoplankton within OMZ [@lavinNovelLineagesProchlorococcus2010; @ulloaCyanobacteriumProchlorococcusHas2021; @partenskyProchlorococcusAdvantagesLimits2010], where they may be net O~2~ consumers [@partenskyComparisonPhotosyntheticPerformances2018].

@@

Climate change is also rapidly changing ocean chemistry.  By the end of this century, surface ocean pH is projected to decline by 0.1 to 0.4  due to projected increases in carbon dioxide concentrations [@garcia-sotoOverviewOceanClimate2021]. Moreover, substantial changes in the global water cycle, leading to extensive changes in worldwide precipitation patterns, are affecting ocean salinity levels on a global scale, and ice melts due to rising temperatures are impacting salinity levels in the Arctic and Northwest Atlantic oceans [@leeRecognizingSalinityThreats2022]. Increasing sea temperatures are also causing decreases in [O~2~] across global oceans [@matearLongtermChangesDissolved2003], particularly toward the poles [@helmObservedDecreasesOxygen2011]. Warmer ocean waters increase stratification, and decrease oxygen solubility at the surface, which in turn decreases oxygen mixing downwards by ocean currents [@garcia-sotoOverviewOceanClimate2021]. Models predict that OMZ in the Pacific and Indian Oceans are expanding [@garcia-sotoOverviewOceanClimate2021; @buseckeDivergingFatesPacific2022]. In spite of these expanding OMZ, Busecke et al. [@buseckeDivergingFatesPacific2022] project that the cores of the OMZ, where the oxygen levels are lowest, may actually contract.

We used the OceansMap Protein Portal (OPP; https://proteinportal.whoi.edu/) [@saitoDevelopmentOceanProtein2021] to analyze the distribution of proteins from clades of *P. marinus* in samples taken across a range of [O~2~] and depth, which in turn correlates to peak light at the site of sampling. We therefore analyzed the growth and physiological responses of representative strains from three clades of *P. marinus* under a matrix of [O~2~], light levels, spectral waveband ranges, and photoperiods, to approximate eco-physiological conditions representative of current and hypothetical future ocean zones. *Prochlorococcus marinus* MED4, a clade HLI strain, was isolated near the ocean surface (5m depth) of the Mediterranean Sea where [O~2~] is near saturation, light levels are high and spectral bias from full solar irradiance is minimal. *Prochlorococcus marinus* SS120, a clade LLIII strain, was isolated from the Sargasso Sea at a depth of 120 m, while *P. marinus* MIT9313, a clade LLIV strain, was isolated from the North Atlantic Gulf Stream at a depth of 135 m. At these depths, light attenuation and spectral shifts occur, resulting in low blue and green light, while [O~2~] varies from near-surface saturation levels to decreased concentrations, but does not necessarily decrease systematically with depth [@billheimerOxygenSeasonalityUtilization2021].

We then analyzed growth rates in terms of cumulative diel Photosynthetically Usable Radiation (PUR). Photosynthetic organisms absorb light energy within the Photosynthetically Active Radiation (PAR) range, 350 to 700 nm, for photosynthesis [@morelAvailableUsableStored1978]. PUR represents the fraction of PAR that is absorbed by the pigments of a given photosynthetic organism [@morelAvailableUsableStored1978], taking into account the specific spectral wavebands these pigments absorb. *Prochlorococcus marinus* light-harvesting complexes, [@chisholmProchlorococcusMarinusNov1992;
@goerickePigmentsProchlorococcusMarinus1992], show an absorption maxima
of 442 nm for divinyl divinylchlorophyll *a* and 478 nm for divinyl chlorophyll b
[@goerickePigmentsProchlorococcusMarinus1992] allowing *P. marinus* to
efficiently harvest blue light in the 400 nm
to 500 nm range [@morelAvailableUsableStored1978] prevailing in deep ocean
habitats, where only blue spectral wavelengths prevail [@holtropVibrationalModesWater2021].  In *P. marinus* small cell diameters, from 0.5 to 0.7 µm [@chisholmProchlorococcusMarinusNov1992], and simple cell structures, minimize the complication of pigment package effect or intracellular self-shading [@morelProchlorococcusSynechococcusComparative1993]. Given the different spectral light regimes typical of the niches of different ecotypes, expressing growth rates in terms of cumulative diel PUR might simplify different photoperiods, spectral bands, and PAR levels into a common parameter, making growth response comparisons across strains and different oxygen levels more accessible.  We also aimed to detect whether growth responses are driven simply by cumulative diel PUR, or whether specific photoperiods, spectral bands or PAR levels have independent, albeit interacting, effects on growth.

@@

Collection and treatment of protein samples were performed following Saito et al. [@saitoMultipleNutrientStresses2014; @saitoNeedlesBlueSea2015]. Briefly, samples of seawater from depths of 20, 40, 50, 60, 70, 80, 90, 120, 150, 200, 250, 300, 380, 400, 500, 550, 600, and 800 m below the ocean surface were pumped through a 0.2 µm filter, preserved in RNAlater and frozen at -80°C until extraction. Proteins were extracted from the filter in an SDS-based detergent, embedded in tube gel, alkylated and reduced prior to in-gel trypsin digestion. Peptide spectra were generated using a Q-Exactive Orbitrap Mass Spectrometer, searched in the SEQUEST CITATIONXXXX and labelled with the most likely protein and species annotation from Uniprot.  Oxygen levels at the location of sampling were recorded.

@@

```{r NATLProchloroPhotosynthDepthO2, fig.cap = NATLProchloroPhotosynthDepthO2_cap , fig.height = 20, fig.width = 10, echo = FALSE}

NATLProchloroPhotosynthDepthO2_cap <- glue("**Ocean detection of *Prochlorococcus marinus* photosynthesis complexes.**", 
                       "Protein detections are plotted vs. O~2~ (µM) (X axis) and depth (m) (Y axis) at sample origin.",
                       "Rows separate data annotated as from *Prochlorococcus marinus* strains MED4 (Clade HLI), SS120 (Clade LLIII) and MIT9313 (Clade LLIV).",
                       "Columns show detections of proteins annotated as  Photosystem II (PSII), Cytochromeb6f complex (Cytb6f), Photosystem I (PSI), ATP Synthase or Ribulose-1,5-bisphosphate oxygenase carboxylase (RUBISCO).",
                       "For comparison culture growth experimental conditions are indicated by horizontal grey lines for depths approximating imposed peak Photosynthetically Active Radiation (µmol photons m^-2^ s^-1^); and vertical grey lines for [O~2~] (µM).",
                       "Point colours identify *Prochlorococcus* clade assignments",
                       "Data from OceanProteinPortal (https://www.oceanproteinportal.org/).",
 .sep = " ")

knitr::include_graphics(file.path(FiguresIn, "NATLProchloroPhotosynthDepthO2.png"))
```

@@

 BluevsRedPurFitsPlots_cap <- glue("**Chlorophyll specific growth rate (d^-1^) vs. cumulative diel Photosynthetic Usable Radiation (PUR, µmol photons m^-2^ d^-1^).**",
                    "Rows show levels of imposed dissolved O~2~ concentrations as 250 µM, 25 µM and 2.5 µM.",
                    "Columns are strains; MED4, SS120 and MIT9313.",
                    "Shapes show the imposed photoperiod (h); 4 h (solid square),  8 h (solid diamond), 12 h (solid circle), 16 h (solid upright triangle).",
                      "Symbol colours show the spectral waveband for growth; 660 nm (red symbols), and 450 nm (blue symbols).",
                      "Large symbols show mean of growth rate from logistic curve fits; small symbols show values for replicate determinations, if any.",
                    "Platt (add Platt citation here) 3 parameter model fit to 660 nm (red lines) and 450 nm (blue lines) growth data for each combination of strain and dissolved oxygen shown with solid lines (significantly different fits, pvalue < 0.05) or dashed lines (not significantly different fits, pvalue > 0.05).",
                       .sep = " ")

@@

```{r NATLProchloroProteinMetabDepthO2, fig.cap = NATLProchloroProteinMetabDepthO2_cap , fig.height = 20, fig.width = 10, echo = FALSE}

NATLProchloroProteinMetabDepthO2_cap <- glue("**Ocean detection of *Prochlorococcus marinus* protein metabolism complexes.**", 
                       "Protein detections are plotted vs. O~2~ (µM) (X axis) and depth (m) (Y axis) at sample origin.",
                       "Rows separate data annotated as from *Prochlorococcus marinus* strains MED4 (Clade HLI), SS120 (Clade LLIII) and MIT9313 (Clade LLIV).",
                       "Columns show detections of proteins annotated as FtsH Protease Complexes (FtsH1, FtsH2, FtsH3) or the Ribosome).",
                       "Culture growth experimental conditions indicated by horizontal grey lines for depths approximating Photosynthetically Active Radiation (µmol photons m^-2^ s^-1^) and vertical grey lines for [O~2~] (µM).",
                       "Point colours identify *Prochlorococcus* clade assignments",
                       "Data from OceanProteinPortal (https://www.oceanproteinportal.org/).",
 .sep = " ")

knitr::include_graphics(file.path(FiguresIn, "NATLProchloroProteinMetabDepthO2.png"))
```

@@

*Prochlorococcus marinus* MED4 has a physiological requirement for more than 4 h of light per day, and thus this strain will not exploit habitats with short photoperiods, typical of winter or light attenuated depths. MED4 is also excluded from the lowest oxygen habitats, equivalent to our 2.5 µM O~2~ experimental conditions. MED4 can, however, grow under OMZ regions with slightly higher [O~2~], as by the growth under our 25 µM O~2~ experiments. Genomic analyses  ( Figure \@ref(fig:ProchlorococcusEnzymeKms))  and previous transcriptional analyses [@bergResponsesPsbAHli2011]  suggests MED4 is excluded from growth below ~ 25 µM O~2~ because it relies upon a ubiquinol oxidase, non-electrogenic, to maintain oxidation/reduction balance in the intersystem electron transport chain.  On the other hand, MED4 shows inducible expression of FtsH isoforms [@bonisteelStrainSpecificDifferences2018], to counter photoinactivation of PSII under higher PAR and [O~2] environments. Photoinactivation, does, however, impose an increased cost of growth upon MED4, since growth under red light, to lower the rate of photoinactivation of PSII [@murphyPhotoinactivationPhotosystemII2017], allows MED4 to achieve faster growth per absorbed photon than growth under red light. TARA Oceans
Project data [@delmontLinkingPangenomesMetagenomes2018] indeed reported presence of *P. marinus* MED4-like genomes at depths ranging from 5 m to 90 m, representing high to low blue light levels, in the Pacific South East Ocean. @delmontLinkingPangenomesMetagenomes2018 did not analyze data from depths beyond the subsurface chlorophyll maximum layer, the layer in the water column where the chlorophyll *a* concentration peaks, nor did they report the [O~2~] at those depths. Our growth findings are also consistent with Figure\@ref(fig:ProchloroPhotosynthDepthO2) showing PSII proteins annotated as MED4, at depths up to 200 meters with O~2~ of 15 µM.
